Language-Based Security for Low-Level MPC

### 2024-05-22 ppdp24/bruteforce.tex
checkpoint
tiny tweaks and a note

In $\mathbb{F}_2$ a brute force strategy for computing
$\progtt(\prog)$ for any $\prog$ is to directly compute
$\runs(\prog)$. By querying $\progtt(\prog)$ we can verify any of the
hyperproperties discussed previously, or other properties such as
perfect secrecy \cite{XXX}.  A basic method to do this is
to calculate a truth table for the given protocol. Since
$\runs(\prog)$ is exponential in
the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
feasible only for smaller protocols. However, to support some
scaling we can efficiently
convert protocols in the passive setting (without $\ttt{assert}$) to
stratified Datalog programs, and then extract $\runs(\prog)$ by
calculating Least Herbrand models by parallelization and/or other HPC
acceleration techniques for logic programs \cite{aspis2018linear}.
The rewriting we describe here is to Datalog with negation, with a
negation-as-failure model, though we can also use techniques in
\cite{sakama2017linear} to eliminate negation from resulting programs.

@@

We immediately note that $\runs(\prog)$ can be obtained by a left-folding
of $\solvealg$ across $\prog$. 
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
    %\cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog
  \end{eqnarray*}
  Then assuming default preprocessing, $\mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog = \runs(\prog)$.
\end{lemma}
However, this method does not take advantage of parallellization,
in that elements of $\runs(\prog)$ can be calculated independently.

>>>>

In the binary field $\mathbb{F}_2$, a brute force strategy for computing
$\progtt(\prog)$ for any $\prog$ is to directly compute
$\runs(\prog)$. By querying $\progtt(\prog)$ we can verify any of the
hyperproperties discussed previously, or other properties such as
perfect secrecy \cite{barthe2019probabilistic}.  A basic method to do this is
to calculate a truth table for the given protocol. Since
$\runs(\prog)$ is exponential in
the size of $\secrets(\prog) \cup \flips(\prog)$, this strategy is
feasible only for smaller protocols. However, to support some
scaling we can efficiently
convert protocols in the passive setting (without $\ttt{assert}$) to
stratified Datalog programs, and then extract $\runs(\prog)$ by
calculating Least Herbrand models by parallelization and/or other HPC
acceleration techniques for logic programs \cite{aspis2018linear}.
The rewriting we describe here is to Datalog with negation, with a
negation-as-failure model, though we can also use techniques in
\cite{sakama2017linear} to eliminate negation from resulting programs.

@@

We immediately note that $\runs(\prog)$ can be obtained by a left-folding
of $\solvealg$ across $\prog$. 
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\xassign{x}{\be}{\cid}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be}{\cid}\ \mathrm{in}\\
      \ \ \{\extend{\store}{x}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{x}{0} \mid \store \in \stores - \stores' \}\end{array}%\\[2mm]
    %\cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog
  \end{eqnarray*}
  Then assuming default preprocessing, $\mathit{foldl}\ {tt}\ \mems(S \cup R)\ \prog = \runs(\prog)$.
\end{lemma}
However, this method does not take advantage of parallellization,
in that elements of $\runs(\prog)$ can be calculated independently.

### 2024-05-22 ppdp24/example-bdoz.tex
checkpoint

In a malicious setting, ``detecting cheating'' by adding
information-theoretic secure MAC codes to shares is a fundamental
approach realized by protocols such as BDOZ \cite{XXX} and SPDZ
\cite{XXX}.  These protocols assume a pre-processing phase that
distributes shares with MAC codes to clients.  This integrates well
with pre-processed Beaver Triples to implement malicious secure, and
relatively efficient, multiplication \cite{XXX}. Recall
that Beaver triples are values $a,b,c$ with $\fcod{a * b} = c$,
unique per multiplication gate, that are secret shared with clients
during pre-processing. Here we consider the 2-party version.

A field value $v$ is secret shared among 2 clients in BDOZ in the same
manner as in GMW.  Each client $\cid$ gets a pair of the form
$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
\emph{global key} $k_\Delta$ is common to all MACs authenticated by
$\cid'$. This is a semi-homomorphic encryption scheme that supports
addition of shares and multiplication of shares by a constant
\cite{XXX}. For more details the reader is referred to \cite{XXX}. We
note that while we restrict values $v$, $m$, and $k$ to the same field
in this presentation for simplicity, in general $m$ and $k$ can be in
extensions of $\mathbb{Z}_p$.

@@

With these conventions, our BDOZ library is defined in Figure \ref{fig-bdoz}.
In $\metaprot$ we represent BDOZ share, MAC pairs as records:
$$
\ttt{\{ share = }v\ttt{;  mac =}\ m \ttt{\}}
$$
and we define $\ttt{macsum}$ for addition of shares,
$\ttt{maccsum}$ for addition of a share and a constant, and
$\ttt{macctimes}$ for multiplication of a share and a constant
in the BDOZ encryption scheme \cite{XXX}. The $\ttt{auth}$
function implements the MAC check as an $\assert$.

@@

The full protocol for malicious secure product of secrets $x$ (that
is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
parties interact in secure opening of $x \fplus a$ and $y + b$,
followed by the non-interactive computation of shares of $x * y$
for secure reveals as per standard protocol
\cite{XXX}. The non-interactive reconstruction of the local authentication
key for both the secure openings and the final reveal is enabled by the
semi-homorphic properties of the BDOZ scheme.

@@

It is straightforward to demonstrate that cheating detection has integrity,
since only the ``passive'' adversary can elicit a response from honest parties. 
\begin{lemma}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

In the case of BDOZ, cheating detection is accomplished by the information-theoretic
security of the encryption scheme\cite{XXX}. Furthermore, the symmetry of
the protocol in Figure \ref{fig-beaver} ensures that both parties will authenticate
shares, so it is robust to corruption of either party.

>>>>

In a malicious setting, ``detecting cheating'' by adding
information-theoretic secure MAC codes to shares is a fundamental
approach realized by protocols such as BDOZ and SPDZ
\cite{SPDZ1,SPDZ2,BDOZ,10.1007/978-3-030-68869-1_3}.  These protocols
assume a pre-processing phase that distributes shares with MAC codes
to clients.  This integrates well with pre-processed Beaver Triples to
implement malicious secure, and relatively efficient, multiplication
\cite{evans2018pragmatic}. Recall that Beaver triples are values $a,b,c$ with
$\fcod{a * b} = c$, unique per multiplication gate, that are secret
shared with clients during pre-processing. Here we consider the
2-party version.

A field value $v$ is secret shared among 2 clients in BDOZ in the same
manner as in GMW.  Each client $\cid$ gets a pair of the form
$(v_\cid,m_\cid)$, where $v_\cid$ is a share of $v$ reconstructed by
addition, i.e., $v = \fcod{v_1 \fplus v_2}$, and $m_\cid$ is a MAC of
$v_\cid$.  More precisely, $m_\cid = k + (k_\Delta * v_\cid)$ where
\emph{keys} $k$ and $k_\Delta$ are known only to $\cid' \ne \cid$ and
$k_\Delta$. The \emph{local key} $k$ is unique per MAC, while the
\emph{global key} $k_\Delta$ is common to all MACs authenticated by
$\cid'$. This is a semi-homomorphic encryption scheme that supports
addition of shares and multiplication of shares by a constant-- for
more details the reader is referred to Orsini
\cite{10.1007/978-3-030-68869-1_3}. We note that while we restrict
values $v$, $m$, and $k$ to the same field in this presentation for
simplicity, in general $m$ and $k$ can be in extensions of
$\mathbb{Z}_p$.

@@

With these conventions, our BDOZ library is defined in Figure \ref{fig-bdoz}.
In $\metaprot$ we represent BDOZ share, MAC pairs as records:
$$
\ttt{\{ share = }v\ttt{;  mac =}\ m \ttt{\}}
$$
and we define $\ttt{macsum}$ for addition of shares,
$\ttt{maccsum}$ for addition of a share and a constant, and
$\ttt{macctimes}$ for multiplication of a share and a constant
in the BDOZ encryption scheme \cite{10.1007/978-3-030-68869-1_3}. The $\ttt{auth}$
function implements the MAC check as an $\assert$.

@@

The full protocol for malicious secure product of secrets $x$ (that
is, $\elab{\secret{x}}{1}$) and $y$ (that is, $\elab{\secret{y}}{2}$)
using Beaver triple $a,b,c$ is defined in Figure \ref{fig-bdoz}. Both
parties interact in secure opening of $x \fplus a$ and $y + b$,
followed by the non-interactive computation of shares of $x * y$ for
secure reveals as per standard protocol
\cite{10.1007/978-3-030-68869-1_3}. The non-interactive reconstruction
of the local authentication key for both the secure openings and the
final reveal is enabled by the semi-homorphic properties of the BDOZ
scheme.

@@

It is straightforward to demonstrate that cheating detection has integrity,
since only the ``passive'' adversary can elicit a response from honest parties. 
\begin{lemma}
  \label{lemma-cheating}
  If cheating is detected in $\prog$, then $\prog$ has integrity.
\end{lemma}

In the case of BDOZ, cheating detection is accomplished by the information-theoretic
security of the encryption scheme \cite{evans2018pragmatic}. Furthermore, the symmetry of
the protocol in Figure \ref{fig-beaver} ensures that both parties will authenticate
shares, so it is robust to corruption of either party.

### 2024-05-22 ppdp24/example-gmw.tex
checkpoint

\paragraph{Oblivious Transfer} A passive secure oblivious transfer (OT) protocol
based on previous work \cite{XXX} can be defined in $\metaprot$,
however this protocol assumes some shared randomness. Alternatively,
a simple passive secure OT can be defined with the addition of
public key cryptography as a primitive. But given the diversity
of approaches to OT, we instead assume that OT is abstract with
respect to its implementation, where calls to OT in $\mathbb{F}_2$
are of the following form-- given a \emph{choice bit}
$\be_1$ provided by a receiver $\cid$, the sender
sends either $\be_2$ or $\be_3$.
$$
\OT{\elab{\be_1}{\cid}}{\be_2}{\be_3}
$$
Critically, the sender learns nothing about $\be_1$ and the
receiver learns nothing about the unselected value, so we interpret
these calls in our implementation in $\mathbb{F}_2$ as follows.
$$
\begin{array}{l}
\solve{\stores}{\OT{\elab{\be_1}{\cid_1}}{\be_2}{\be_3}}{\cid_2} = \\
\qquad ((\solve{\stores}{\be_1}{\cid_1}) \cap 
(\solve{\stores}{\be_2}{\cid_2})) \cup \\
\qquad ((\stores - (\solve{\stores}{\be_1}{\cid_1})) \cap
(\solve{\stores}{\be_3}{\cid_2})
\end{array}
$$

@@

As discussed above and in related work \cite{XXX}, probabilistic
separation conditional on certain variables-- e.g., secret inputs or
public outputs-- is a key mechanism for reasoning about MPC protocol
security. Following \cite{XXX}, we define a conditional separation
relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that under the
condition of all value assignments for $X_2$ and $X_3$ are separable
under $\pmf$ on the condition of any value assignment of $X_1$-- i.e.,
conditionally on any $\store \in \mems(X_1)$. Another key concept
needed especially for reasoning about circuits is conditional
determinism. For example, if $\macgv{z}$ is an output of an internal
gate, it will definitely be computed using random variables, however,
it \emph{should} be determistic under any set of input secrets $S$,
since we assume that $\idealf$ is deterministic. Conditional
uniformity is also important, since the gradual release property of
many protocols means that messages appear in a uniform distribution to
the adversary.
\begin{definition}[Conditioning Properties] 
  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
  we write:
  \begin{itemize}
  \item $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$ we have
    $\pmf(\store|\store') = \pmf(\store_{X_2}|\store') *  \pmf(\store_{X_3}|\store')$.
  \item $\conddetx{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ there exists 
    $\store \in \mems(X_2)$ such that $\pmf(\store|\store') = 1$.
  \item $\condunix{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ and
    $\store \in \mems(X_2)$ we have
    $\pmf(\store|\store') = 1/p^{|X_2|}$.
  \end{itemize}
\end{definition}

>>>>

\paragraph{Oblivious Transfer} A passive secure oblivious transfer (OT) protocol
based on previous work \cite{barthe2019probabilistic} can be defined in $\metaprot$,
however this protocol assumes some shared randomness. Alternatively,
a simple passive secure OT can be defined with the addition of
public key cryptography as a primitive. But given the diversity
of approaches to OT, we instead assume that OT is abstract with
respect to its implementation, where calls to OT in $\mathbb{F}_2$
are of the following form-- given a \emph{choice bit}
$\be_1$ provided by a receiver $\cid$, the sender
sends either $\be_2$ or $\be_3$.
$$
\OT{\elab{\be_1}{\cid}}{\be_2}{\be_3}
$$
Critically, the sender learns nothing about $\be_1$ and the
receiver learns nothing about the unselected value, so we interpret
these calls in our implementation in $\mathbb{F}_2$ as follows.
$$
\begin{array}{l}
\solve{\stores}{\OT{\elab{\be_1}{\cid_1}}{\be_2}{\be_3}}{\cid_2} = \\
\qquad ((\solve{\stores}{\be_1}{\cid_1}) \cap 
(\solve{\stores}{\be_2}{\cid_2})) \cup \\
\qquad ((\stores - (\solve{\stores}{\be_1}{\cid_1})) \cap
(\solve{\stores}{\be_3}{\cid_2})
\end{array}
$$

@@

As discussed above and in related work \cite{8429300}, probabilistic
separation conditional on certain variables-- e.g., secret inputs or
public outputs-- is a key mechanism for reasoning about MPC protocol
security. Following \cite{barthe2019probabilistic}, we define a conditional separation
relation $\condsep{\pmf}{X_1}{X_2}{X_3}$ to mean that under the
condition of all value assignments for $X_2$ and $X_3$ are separable
under $\pmf$ on the condition of any value assignment of $X_1$-- i.e.,
conditionally on any $\store \in \mems(X_1)$. Another key concept
needed especially for reasoning about circuits is conditional
determinism. For example, if $\macgv{z}$ is an output of an internal
gate, it will definitely be computed using random variables, however,
it \emph{should} be determistic under any set of input secrets $S$,
since we assume that $\idealf$ is deterministic. Conditional
uniformity is also important, since the gradual release property of
many protocols means that messages appear in a uniform distribution to
the adversary.
\begin{definition}[Conditioning Properties] 
  Give $\pmf$ and $X_1,X_1,X_3 \subseteq \dom{pmf}$, 
  we write:
  \begin{itemize}
  \item $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store' \in \mems(X_1)$ and $\store \in \mems(X_2 \cup X_3)$ we have
    $\pmf(\store|\store') = \pmf(\store_{X_2}|\store') *  \pmf(\store_{X_3}|\store')$.
  \item $\conddetx{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ there exists 
    $\store \in \mems(X_2)$ such that $\pmf(\store|\store') = 1$.
  \item $\condunix{\pmf}{X_1}{X_2}$ iff for all
    $\store' \in \mems(X_1)$ and
    $\store \in \mems(X_2)$ we have
    $\pmf(\store|\store') = 1/p^{|X_2|}$.
  \end{itemize}
\end{definition}

### 2024-05-22 ppdp24/metalang.tex
checkpoint

The $\fedprot$ language is low-level and does not include abstractions for
defining composable elements. So in this Section we introduce the $\metaprot$
language which that includes structured data and function definitions
for defining composable protocol elements at a higher level of abstraction. 
The $\metaprot$ language is a \emph{metalanguage}, where 
$\fedprot$ protocols are the residuum of computation. In addition to these
declarative benefits of $\metaprot$, component
definitions support compositional verification of larger programs as we will
discuss with examples in Section
\ref{section-examples}.

>>>>

The $\fedprot$ language is low-level and does not include abstractions
for defining composable elements. So in this Section we introduce the
$\metaprot$ language which that includes structured data and function
definitions for defining composable protocol elements at a higher
level of abstraction.  The $\metaprot$ language is a
\emph{metalanguage}, where $\fedprot$ protocols are the residuum of
computation. In addition to these declarative benefits of $\metaprot$,
component definitions support compositional verification of larger
prortocols as we will discuss with examples in Sections
\ref{section-example-gmw} and \ref{section-example-bdoz}.
